---
title: "pins: manage, discover and share datasets in R."
output:
  github_document:
    fig_width: 9
    fig_height: 5
---

[![Build Status](https://travis-ci.org/javierluraschi/pins.svg?branch=master)](https://travis-ci.org/javierluraschi/pins) [![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/pins)](https://cran.r-project.org/package=pins)

- **Manage** your personal datasets by pinning and retrieving them with `pin()`.
- **Discover** new datasets from R packages, online and in your organization using `find_pin()`.
- **Share** existing **datasets** online and within your organization using `publish_pin()`.
- **Extend** storage locations using **boards** through `use_board()`, you decide where your data lives.

## Installation

You can install `pins` using the `remotes` package:

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("javierluraschi/pins")
```

## Personal datasets

You can track personal dataset by pinning them as follows:

```{r}
library(dplyr)
library(pins)

iris %>%
  filter(Sepal.Width < 3, Petal.Width < 1) %>%
  pin("iris-small-width", "A subset of 'iris' with only small widths.")
```

You can then use this dataset as,

```{r}
pin("iris-small-width")
```

A pin is a tool to help you organize content, not the content itself. Therefore, you should not use a pin to store your findings; instead, you should still persist files or check-in reproducible code into GitHub. Pins are here to help retrieve and find datasets.

The motivation behind pinning is to allow you to easily fetch results from past data analysis sessions. This can be useful after tidying your data, since once a dataset is tidy you are likely to reuse this several times and. You might also have a past analysis in GitHub, but you might not want to clone, install dependencies and rerun your code just to access your dataset, that's another good case for using a pin. Another use case is to cross-join between datasets to analyse across multiple projects or help you remember which datasets you've used in the past.

You can find previous datasets using `find_pin()`:

```{r}
find_pin()
```

## Shared datasets

`pins` supports shared storage locations using boards. A board is a remote location for you to share datasets with your team privately, or with the world, publicly. Use `use_board()` to choose a board, currently only databases are supported; however, `pins` provide an extensible API you can use to store pins anywhere.

### Databases

First, you will need to create a database connection using `DBI`, it is outside the scope of this tutorial to teach you how to connect to a database using `DBI`; however, there are many resources available online.

For instance, you can use a `bigrquery` connection to define our database-backed board,

```{r echo=FALSE}
bq_project <- Sys.getenv("BQ_PROJECT")
bq_dataset <- Sys.getenv("BQ_DATASET")
```
```{r}
use_board("database", DBI::dbConnect(bigrquery::bigquery(), project = bq_project, dataset = bq_dataset))
```

Which we can also use to pin a dataset,

```{r}
pin(iris, "iris", "The entire 'iris' dataset.")
```

Find pins,

```{r}
find_pin()
```

And retrieve datasets,

```{r}
pin("iris")
```
