---
title: "Understanding Boards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Understanding Boards}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, echo = FALSE, message = FALSE}
library(nomnoml)
```

A board is an storage location to store pins for personal use, with your team, or publicly with the world. A board defines four operations to manage pins in any storage system: create, find, retrieve and remove. A board can be read-only if it only supports the retrieve and find operations.

```{nomnoml echo=FALSE}
[Pin]->[Create]
[Create]->[Board]
[Pin]<-[Find]
[Find]<-[Board]
[Pin]<-[Retrieve]
[Retrieve]<-[Board]
[Pin]->[Remove]
[Remove]->[Board]
```

## Files

The defaul board in the `pins` package is a files-based; meaning, data is stored in the file system assigned to your R session. When using this from your personal computer, this defaults to a `pins` folder under the default home folder for your user.

```{nomnoml echo=FALSE}
[R|
  [<note>library(pins)
    pin(iris, "iris")]
]-[File System|
  [<note>~/documents/
   ~/images/
   ~/pins/files/pins.yml
   ~/pins/files/iris.Rds
  ]
]
```

## Database

```{nomnoml echo=FALSE}
[R|
   [<note>library(pins)
    use_board("database", DBI::dbConnect())
    pin(iris, "iris")]
]-[<database>Database|
  [<note>db.pins
  db.pins_iris]]
```

## RStudio

```{nomnoml echo=FALSE}
[R|
   [<note>library(pins)
    use_board("rstudio")
    pin(iris, "iris")]
]-[RStudio Connect|
  [<note>iris]]
```

When using multiple publishing servers, you can specify an specific server through `use_board("rstudio", "<server-name>")`.

## Kaggle

```{nomnoml echo=FALSE}
[R|
   [<note>library(pins)
    use_board("kaggle")
    pin(iris, "iris")]
]-[Kaggle|
  [<note>iris]]
```

## Other

If the boards that the `pins` package provides are not sufficient, say, you might want to store your pins in Google Drive; you can still implement custom boards by implementing a few S3 methods. It is advisable to create an R package that provides these methods, but you can also define them in a plain R script.

The following example creatse an in-memory board; probably not suitable for real use cases since all your pins would be lost whenever the R session restarts; however, it's a simple example that you can use to start developing custom boards:

```{r}
board_initialize.memory <- function(board, ...) {
  as.environment(list(data = data.frame(name = c(), description = c(), type = c(), metadata = c())))
}

pin_create.memory <- function(board, x, name, description, type, metadata) {
  board$data <- rbind(board$data,
    data.frame(name = name, description = description, type = type, metadata = metadata)
  )
}

pin_find.memory <- function(board, text) {
  if (is.null(text)) return(.memory_board)
  board$data[grepl(text, board$data$description) | board$data$name == text,]
}

pin_retrieve.memory <- function(board, name) {
  board$data[board$data$name == name]
}

pin_remove.memory <- function(board, name) {
  board$data <- board$data[board$data$name != name]
}

board_info.memory <- function(board) {
}
```
