---
title: "Managing custom formats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Managing custom formats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = rlang::is_installed("arrow")
)
```

The pins package provides a robust set of functions to read and write standard types of files using standard tools, e.g. CSV files using `read.csv()` and `write.csv()`.
However, from time to time, you may wish read or write using other tools. You may want to read and write:

 - CSV files using readr or vroom
 - Arrow files without using compression
 - Parquet files
 - Whole directories that are archived/zipped

An escape hatch for a customized approach is provided: `pin_upload()` and `pin_download()`. 
The goal of this vignette is to show how you can incorporate these into your workflow.

Two points to keep in mind:

 - `pin_upload()` takes a vector of `paths` to local files.
 - `pin_download()` returns a vector of `paths` to local files.

We'll follow an example where we write and read uncompressed Arrow files, starting by creating a temporary board:

```{r setup}
library(pins)

board <- board_temp()
```

## Upload a one-off file

If you are writing a one-off file, you can do everything directly:

```{r}
pin_name <- "mtcars-arrow"

# file name will be `mtcars-arrow.arrow`
path <- fs::path_temp(fs::path_ext_set(pin_name, "arrow"))

arrow::write_feather(mtcars, path, compression = "uncompressed")

pin_upload(board, paths = path, name = pin_name)
```

Reading from the downloaded pin is straightforward; `pin_download()` returns a local path that can be piped to `arrow::read_feather()`:

```{r}
mtcars_download <- 
  pin_download(board, pin_name) %>%
  arrow::read_feather()

head(mtcars_download)
```

## Function to manage uploading

If you want to write more than one custom file of a certain type, or using a certain tool, you might consider writing a helper function:

```{r}
pin_upload_arrow <- function(board, x, name, ...) {
  # path deleted when `pin_upload_arrow()` exits
  path <- fs::path_temp(fs::path_ext_set(name, "arrow"))
  withr::defer(fs::file_delete(path))
 
  # custom writer
  arrow::write_feather(x, path, compression = "uncompressed")

  pin_upload(board, paths = path, name = name, ...) 
}
```

This helper function is designed to work like `pin_write()`:

```{r}
pin_upload_arrow(board, x = mtcars, name = "mtcars-arrow2")
```

As before, you can pipe the result of `pin_download()` to your reader function:

```{r}
pin_download(board, name = "mtcars-arrow2") %>%
  arrow::read_feather() %>%
  head()
```

## Another example function

If you want to use this same approach to [archive](https://archive.r-lib.org/) and pin a whole directory, you can write a helper function like:

```{r}
pin_upload_archive <- function(board, dir, name, ...) {
  path <- fs::path_temp(fs::path_ext_set(name, "tar.gz"))
  withr::defer(fs::file_delete(path))
  archive::archive_write_dir(path, dir)
  pin_upload(b, paths = path, name = name, ...)
}
```

You can download the compressed archive via `pin_download(board, name)` and then pipe that path straight to `archive::archive_extract()` to extract your archive in a new diretory.
