---
title: "Upgrading to pins 1.0.0"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Upgrading to pins 1.0.0}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

pins 1.0.0 introduced a completely new API designed to be more explicit and less magical.
While the legacy API will continue to be supported for some time, it will not gain any new features, so it's good to plan to switch to the new interface.

This vignette shows a couple of examples converting legacy code to the modern API, then shows the equivalences between the legacy and modern function names.

```{r setup}
library(pins)
```

## Examples

A simple example of the legacy API looks something like this:

```{r}
board_register_local("vignette", tempfile())

pin(head(mtcars), "mtcars", board = "vignette")
pin_get("mtcars", board = "vignette")
```

There are three major changes:

-   You create an explicit board object which you use instead of the name.
-   You use `pin_read()` and `pin_write()` instead of `pin_get()` and `pin()`.
-   The board object is the always the first argument.

So the code above translates to:

```{r}
board <- board_local()

board %>% pin_write(head(mtcars), "mtcars")
board %>% pin_read("mtcars")
```

### Pinning files

Another way to use `pin()` is with a path to a file:

```{r}
path <- tempfile()
writeLines(letters, path)

pin(path, "alphabet", board = "vignette")
pin_get("alphabet", board = "vignette")
```

pins 1.0.0 clearly separates the two cases of pin an object and pinning a file, so instead of `pin_write()` and `pin_read()` you need to `pin_upload()` and `pin_download()`:

```{r}
board %>% pin_upload(path, "alphabet")
board %>% pin_download("alphabet")
```

### Pinning a url

Finally, you can `pin()` a url to automatically re-download it when it changes:

```{r}
base <- "https://raw.githubusercontent.com/rstudio/pins/master/tests/testthat/"

(pin(paste0(base, "pin-files/first.txt"), board = "vignette"))
```

(It's a little confusing that this uses `pin()` rather than `pin_get()`.)

This now needs to be made explicit with the new `board_url()`, and since this returns a path, not a file, you need to use `pin_download()`:

```{r}
board_github <- board_url(c(
  raw = paste0(base, "pin-files/first.txt")
))
board_github %>% pin_download("raw")
```

### Implicit board

It's also possible to use `pin()` and `pin_get()` without an explicit board argument, in which case it automatically uses a local board:

```{r}
pin(data.frame(x = 1:3), "test-data")
pin_get("test-data")
```

To convert this, you just need to create an explicit board:

```{r}
board <- board_local()

board %>% pin_write(data.frame(x = 1:3), "test-data")
board %>% pin_read("test-data")
```

## Equivalents

### Board functions

| Legacy API                   | Modern API          |
|------------------------------|---------------------|
| `board_register_azure()`     |                     |
| `board_register_datatxt()`   |                     |
| `board_register_dospace()`   |                     |
| `board_register_gcloud()`    |                     |
| `board_register_github()`    |                     |
| `board_register_local()`     | `board_local()`     |
| `board_register_kaggle()`    |                     |
| `board_register_rsconnect()` | `board_rsconnect()` |
| `board_register_s3()`        | `board_s3()`        |
| `pin()` with a URL           | `board_url()`       |
|                              | `board_folder()`    |

Future releases will add support for additional boards based on user feedback.

### Pin functions

| Legacy API       | Modern API                                        |
|------------------|---------------------------------------------------|
| `board_browse()` | `pin_browse()`                                    |
| `pin()`          | `pin_write()` / `pin_upload()`                    |
| `pin_get()`      | `pin_read()` / `pin_upload()`                     |
| `pin_find()`     | `pin_search()`                                    |
| `pin_info()`     | `pin_meta()`                                      |
| `pin_reactive()` | `pin_reactive_read()` / `pin_reactive_download()` |
| `pin_remove()`   | `pin_delete()`                                    |
