---
title: "Get started with pins"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with pins}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
) 

# Fake testing environment to get consistent hashes
Sys.setenv("TESTTHAT" = "true")
options(pins.quiet = FALSE)
```

The pins package helps you publish data sets, models, and other R objects, making it easy to share them across projects and with your colleagues.
You can pin objects to a variety of "boards", including local folders (to share on a networked drive or with dropbox), RStudio connect, Amazon S3, and more.
This vignette will introduce you to the basics of pins.

```{r}
library(pins)
```

## Getting started

Every pin lives in a pin board, so you have to start by creating a pin board.
In this vignette I'm going to use a temporary board which is automatically deleted when your R session is over:

```{r}
board <- board_temp()
```

But in real-life, you'd pick another board depending on where you want the pins data to live.
Here are a few options:

```{r, eval = FALSE}
board <- board_local() # share data across R sessions on the same computer
board <- board_folder("~/Dropbox") # share data with others using dropbox
board <- board_folder("Z:\\my-team\pins") # share data using a shared network drive
board <- board_rsconnect() # share data with RStudio connect
```

## Reading and writing data

Once you have a pin board, you can save data to it with `pin_write()`:

```{r}
mtcars <- tibble::as_tibble(mtcars)
board %>% pin_write(mtcars, "mtcars")
```

The first argument is the object to save (usually a data frame, but it can be any R object), and the second argument defines the "name" of pin.
The name is important and is basically equivalent to a file name: you'll use it when you later want to read the data from the pin.
The only rule for a pin name is that it can't contain slashes.

As you can see, pins has chosen to save this data to a `.rds` file.
But you can choose from a variety of other options depending on your goals:

-   `type = "rds"` uses `writeRDS()` to save a binary R data file. It can save any R object but it's only readable from R, not other languages.
-   `type = "csv"` uses `write.csv()` to write a `.csv` file. CSVs can read by any application, but only support a simple columns (e.g. numbers, strings, dates) and are largely large and slow.
-   `type = "arrow"` uses `arrow::write_feather()` to create an arrow/feather file. [Arrow](https://arrow.apache.org) is a modern, language-independent, high-performance, file format for design for data science. Not every tool can read arrow files, but support is growing rapidly.
-   `type = "json"` uses `jsonlite::write_json()` to create a `.json` file. Pretty much every programming language can read json files, but they only work well for nested lists.

In a future session, you can read that data back with `pin_read()`:

```{r}
board %>% pin_read("mtcars")
```

You don't need to supply the file type when reading data from a pin because pins automatically stores that in the metadata, the topic of the next section.

Note that when the data lives elsewhere, pins takes care of downloading and caching so that it's only re-downloaded when needed.
That said, most boards transmit pins over HTTP, and this is going to be slow and possibly unreliable for very large pins.
As a general rule of thumb, we don't recommend using pins with files over 500 MB.
If you find yourself routinely pinning data larger that this, you might need to reconsider your data engineering pipeline.

## Metadata

Every pin is accompanied by some metadata that you can access with `pin_meta()`:

```{r}
board %>% pin_meta("mtcars")
```

This shows you the metadata that's generated by default.
This includes:

-   `description`, a brief textual description of the dataset.

-   `created`, the date-time when the pin was created.

-   `file_size`, the size (in bytes) of the underlying file.

-   `pin_hash`, a unique hash that you can supply to `pin_read()` to ensure that you're reading exactly the data that you expect.

When creating the pin, you can override the default description and provide additional metadata that is stored with the data:

```{r}
board %>% pin_write("mtcars", 
  desc = "Data extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).",
  metadata = list(
    source = "Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391–411."
  )
)
board %>% pin_meta("mtcars")
```

While we'll do our best to keep the automatically generated metadata consistent over time, I'd recommend manually capturing anything you really care about in `metadata`.

## Versioning

In many situations it's useful to version pins, so that writing to an existing pin does not replace the existing data, but instead adds a new copy.
There are two ways to turn versioning on:

-   When you create a board you can turn versioning on for every pin in that board:

    ```{r, eval = FALSE}
    board2 <- board_temp(versioned = TRUE)
    ```

-   When you write a pin, you can specifically request that versioning be turned on for that pin:

    ```{r, eval = FALSE}
    board2 <- board_temp()
    board2 %>% pin_write(mtcars, versioned = TRUE)
    ```

Some boards, like `board_rsconnect()` have versioning on by default, because it's automatically supported by the underlying storage engine.
In most cases, however, you'll need to explicitly enable it.

Once you have turned versioning on, every `pin_write()` will create a new version:

```{r}
board2 <- board_temp(versioned = TRUE)

board2 %>% pin_write(1:5, name = "x", type = "rds")
board2 %>% pin_write(2:6, name = "x", type = "rds")
board2 %>% pin_write(3:7, name = "x", type = "rds")
```

You can list all the available versions with `pin_versions()`:

```{r}
board2 %>% pin_versions("x")
```

(Pins does not currently provide any tool to remove old versions, but this is likely to be implemented in the future.)

By default, `pin_read()` will return the most recent version:

```{r}
board2 %>% pin_read("x")
```

But you can request an older version by supplying the `version` argument:

```{r}
board2 %>% pin_read("x", version = "ab444e71e875d63b")
```

## Reading and writing files

So far we've focussed on `pin_write()` and `pin_read()` which work with R objects.
pins also provides the lower-level `pin_upload()` and `pin_download()` which work with files on disk.
You can use them to share data that's otherwise unsupported by pins.

`pin_upload()` works like `pin_write()` but instead of an R object you give it a vector of paths.
I'll start by creating a few files in the temp directory:

```{r}
paths <- file.path(tempdir(), c("mtcars.csv", "alphabet.txt"))
write.csv(mtcars, paths[[1]])
writeLines(letters, paths[[2]])
```

Now I can upload those to the board:

```{r}
board %>% pin_upload(paths, "example")
```

`pin_download()` returns a vector of paths:

```{r}
board %>% pin_download("example")
```

It's now your job to handle them.
You should treat these paths as internal implementation details --- never modify them and never save them for use outside of pins.

Note that you can't `pin_read()` something you pinned with `pin_upload()`:

```{r, error = TRUE}
board %>% pin_read("example")
```

But you can `pin_download()` something that you've pinned with `pin_write()`:

```{r}
board %>% pin_download("mtcars")
```

## Caching

The primary purpose of pins is to make it easy to share data.
But pins is also designed to help you spend as little time as possible downloading data.
`pin_read()` and `pin_download()` automatically cache remote pins: they maintain a local copy of the data (so it's fast) but always check that it's up to date (so your analysis doesn't use stale data).

Wouldn't it be nice if you could take advantage of this feature for any dataset on the internet?
That's the idea behind `board_url()` --- you can assemble your own board from datasets, wherever they live on the internet.
For example, this code create a board containing a single pin, `penguins`, that refers to some data I found on GitHub:

```{r}
my_data <- board_url(c(
  "penguins" = "https://raw.githubusercontent.com/allisonhorst/palmerpenguins/master/inst/extdata/penguins_raw.csv"
))
```

You can read this data by combining `pin_download()` with `read.csv()`[^1]:

[^1]: Here I'm using `read.csv()` to the reduce the dependencies of the pins package.
    For real code I'd recommend using `data.table::fread()` or `readr::read_csv().`

```{r}
my_data %>%
  pin_download("penguins") %>% 
  read.csv(check.names = FALSE) %>% 
  tibble::as_tibble()
```

`board_url()` requires a bit of work compared to using `download.file()` or similar but it has a big payoff: the data is cached locally (so it's fast to read) and will be re-downloaded when it changes (so it's never stale).
