---
title: "Use Cases"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("utils.R")
```

There are many interesting use cases available for `pins`, which we present next.

For instance, after cleaning up your data, you can consider letting others [reuse your tidy datasets](use-cases/reusing-tidy-datasets.html). Once your data is in a shared location, you might want to consider to [schedule dataset updates](use-cases/schedule-dataset-updates.html) automatically. It is then natural to attempt to pipe multiple scheduled datasets to [create simple data pipelines](use-cases/create-simple-data-pipelines.html) or [update plumber and shiny apps](use-cases/update-plumber-and-shiny-apps.html).

For ease of use, the use cases are presented using [RStudio Connect](https://rstudio.com/products/connect/); however, similar workflows can be built in, say GitHub, by making use of [Travis](https://travis-ci.org), [AppVeyor](http://appveyor.com) or [GitHub Actions](https://github.com/features/actions).

```{r}
pins_site_gallery(list(
  list(image       = "images/use-cases-reuse-tidy-datasets.png",
       title       = "Reuse Tidy Datasets",
       href        = "use-cases/reuse-tidy-datasets.html",
       description = "Doing data analysis usually requires tidying your data, a tidy dataset is something worth considering sharing with others or your future self. You can easily incorporate 'pins' in your data analysis workflows by sharing your tidy dataset with others in any of the supported boards (GitHub, Kaggle, RStudio Connect, etc).",
       app         = "https://beta.rstudioconnect.com/connect/#/apps/6522/access"),
   list(image      = "images/use-cases-schedule-dataset-updates.png",
       title       = "Schedule Dataset Updates",
       href        = "use-cases/schedule-dataset-updates.html",
       description = "Once a dataset is shared with as a pin, it is natural to consider automating the process of updating it. This is especially relevant for datasets that change frequently. You can use automated tools like Travis, AppVeyor, GitHub Actions or RStudio Connect in combination with R and the pins package to keep these dataset up to date.",
       app         = "https://beta.rstudioconnect.com/connect/#/apps/6446/access"),
   list(image       = "images/use-cases-reuse-tidy-datasets.png",
       title       = "Create Simple Data Pipelines",
       href        = "use-cases/create-simple-data-pipelines.html",
       description = "",
       app         = ""),
   list(image       = "images/use-cases-reuse-tidy-datasets.png",
       title       = "Update Plumber and Shiny Apps",
       href        = "use-cases/update-plumber-and-shiny-apps.html",
       description = "",
       app         = "")
))
```
