---
title: "Using web-hosted boards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using web-hosted boards}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Pins offers a lot of ways to manage back-and-forth collaboration using, for example, `board_s3()`.
The goal of this vignette is to show how a `board_url()` can bring your pins to a wider audience.

A couple of things to keep in mind:

 - For consumers, `board_url()` offers read-only collaboration.
 - For publishers, a helper function is provided, `write_board_manifest()`.
   A manfiest contains a list of pins and versions; this function lets
   you provide consumers with the same access to the board as you.
   
The rest of the vignette will go into greater detail on these points, providing examples.

## Publishing

```{r setup}
library(pins)
```

```{r pin-mtcars}
board <- board_temp()

board %>% pin_write(mtcars, type = "json")
```

```{r board-pin-list}
board %>% pin_list()

board %>% pin_meta("mtcars")

board %>% pin_versions("mtcars")
```

```{r}
board %>% write_board_manifest()
```


```yaml
`r paste(readLines(fs::path(board$path, "_pins.yaml")), collapse = "\n")`
```

```{r board-serve}
#| echo: false
board_server <- webfakes::new_app()
board_server$use(webfakes::mw_static(root = board$path))
board_process <- webfakes::new_app_process(board_server)

web_board <- board_url(board_process$url())
```

## Consuming

```{r}
#| message: false
web_board %>% pin_list()

web_board %>% pin_versions("mtcars")

web_board %>% pin_meta("mtcars")

web_board %>% pin_read("mtcars") %>% head()
```

## Publishing platforms

### Quarto

### pkgdown

### S3
