---
title: "Using web-hosted boards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using web-hosted boards}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = rlang::is_installed("webfakes")
)
```

The pins package supports back-and-forth collaboration using, for example, `board_s3()`.
The goal of this vignette is to show how to publish a board of pins to a website, bringing your pins to a wider audience.

A couple of things to keep in mind:

 - For consumers, `board_url()` offers read-only collaboration.
 - For publishers, a helper function is provided, `write_board_manifest()`.
   A manifest contains a list of pins and versions, enabling
   a `board_url()` to read like a `board_s3()` or `board_folder()`.

## Publishing

The general steps for publishing a board that can be read using `board_url()` are:

 - create a board (if needed)
 - write pins to board
 - write a manifest file, using `write_board_manifest()`
 - publish website

The first and last steps depend on how you create and deploy your board; we discuss these in the [Publishing platforms] section. Regardless of platform, you'll write the pins and the manifest the same way.

For this first demonstration, we'll start by creating a `board_temp()`, and finish by "magically" serving the board.

```{r setup}
library(pins)
```

In practice, you might use a `board_folder()` in a project folder, or perhaps a `board_s3()`.
For the purpose of this vignette, we create a temporary board.

```{r pin-mtcars}
board <- board_temp(versioned = TRUE)
```

Let's make the `mtcars` dataset available as a JSON file:

```{r}
board %>% pin_write(mtcars, type = "json")
```

Let's make a new version of this pinned data by adding a new column: `lper100km`, consumption in liters-per-100-km.
This could make our data friendlier to folks outside the United States.


```{r}
#| echo: false

# we need a short delay here to keep the versions in the right order
Sys.sleep(2)
```

```{r metric}
mtcars_metric <- mtcars
mtcars_metric$lper100km <- 235.215 / mtcars$mpg

board %>% pin_write(mtcars_metric, name = "mtcars", type = "json")
```

We make a quick check of our board to ensure we have one pin, named `"mtcars"`, with two versions:

```{r board-pin-list}
board %>% pin_list()

board %>% pin_versions("mtcars")
```

Because a `board_url()` is consumed over the web, it doesn't have access to a file system the way, for example, a `board_folder()` has; our workaround is to create a manifest file.
When a `board_url()` is set up by a consumer for reading, the function uses this file to discover the pins and their versions.
The manifest file is the key to `board_url()`'s ability to discover pins as if it were a file-system-based board.
 
After writing pins, but before publishing, call `write_board_manifest()`:

```{r}
board %>% write_board_manifest()
```

We confirm that there is a file called `_pins.yaml`:

```{r}
withr::with_dir(board$path, fs::dir_ls())
```

We can inspect its contents to see each pin in the board, and each version of each pin:

```yaml
`r if (rlang::is_installed("webfakes")) paste(readLines(fs::path(board$path, "_pins.yaml")), collapse = "\n")`
```

At this point, we would publish the folder containing the board as a part of a web site.
For the purpose of this vignette, we pretend that we have served the folder from our fake website, `https://not.real.website.co/pins/`.

```{r board-serve}
#| echo: false
board_server <- webfakes::new_app()
board_server$use(webfakes::mw_static(root = board$path))
board_process <- webfakes::new_app_process(board_server)

web_board <- board_url(board_process$url())
```

## Consuming

With an up-to-date manifest file, a `board_url()` can behave as a read-only version of a `board_folder()`.
Let's create a `board_url()` using our fake URL:

```{r}
#| eval: false
web_board <- board_url("https://not.real.website.co/pins/")
```

The `board_url()` function reads the manifest file to discover the pins and versions:

```{r}
#| message: false
web_board %>% pin_list()

versions <- web_board %>% pin_versions("mtcars") %>% print()
```

We can read the most-recent version of the `"mtcars"` pin:

```{r}
#| message: false
web_board %>% pin_read("mtcars") %>% head()
```

We can also read the first version:

```{r}
#| message: false
web_board %>% pin_read("mtcars", version = versions$version[[1]]) %>% head()
```

The next section shows some ways you can integrate web-based boards into your workflows.

## Publishing platforms

The goal of this section is to illustrate, for a few platforms, the particular way to publish a board as a part of a website.

### Quarto

Using Quarto, you can [publish](https://quarto.org/docs/publishing/) to GitHub Pages, Quarto Pub, and Netlify, among others.

Let's say you set up a board using a directory called `"pins-board"`; of course, you may wish to use a different name.
Here's how to set up your Quarto project:

1. In your code, create a `board_folder("pins-board")` in the root directory:
  - Write to the board using `pin_write()`.
  - At the end, call `write_board_manifest()`.

2. Add the board folder as a [Quarto site-resource](https://quarto.org/docs/websites/website-tools.html#site-resources):

  - At the top level of your `_quarto.yml` file:
  
    ```yml
    resources:
      - pins-board
    ```

  - You may wish to add `/pins-board/` to your project's `.gitignore` file.

Here's an example of a [Quarto repository](https://github.com/ijlyttle/quarto-pins) and its [published website](https://ijlyttle.github.io/quarto-pins/).

### pkgdown

Pins offers another way for R-package developers to share data associated with a package. 
Publishing a package dataset as a pin can let you extend your data's "audience" to Python users.

Using [pkgdown](https://pkgdown.r-lib.org/articles/customise.html#additional-html-and-files), any files you save in the directory `pkgdown/assets/` will be copied to the website's root directory when `pkgdown::build_site()` is run.

The [R Packages book](https://r-pkgs.org/data.html#sec-data-data-raw) suggests using a folder called `data-raw` for working with datasets; this can be adapted to use pins. You would start with `usethis::use_data_raw()`.

In your data-raw file, you would wrangle and clean your datasets, then `usethis::use_data()` to create the datasets for your package.
To offer such datasets on a web-based board, in your data-raw file you would:

  - Create a board: `board_folder(here::here("pkgdown/assets/pins-board"))`; you might use a different name than `"pins-board"`.
  - Write your datasets to the board using `pin_write()`.
  - At the end, call `write_board_manifest()`.

The R Packages book offers this observation on CRAN and package data:

> Generally, package data should be smaller than a megabyte - if it’s larger you’ll need to argue for an exemption.

Publishing a board on your pkgdown site provides a way to offer extended versions of your data.

Here's an example of a [toy package](https://github.com/ijlyttle/pkgpins) and its [pkgdown site](https://ijlyttle.github.io/pkgpins/).

### S3

My understanding is that if you create a `board_s3()`, then allow anonymous access to the S3 bucket, things should "just work". Providing, of course, that you `write_board_manifest()` in a timely manner.

