---
title: "RStudio Connect -- World News"
---

## News Pin

The following chunks creates a list with the latest world news from the BBC:

```{r retrieve-world-news}
library(xml2)

world_news <- data.frame(
  title = xml_text(
    xml_find_all(
      read_xml("http://feeds.bbci.co.uk/news/rss.xml"),
      "///item/title/node()"
    )
  )
)
```

We can then run this under RStudio Connecet on, say, a daily schedule; for this to work, you first need to register the `rstudio`:

```{r}
pins::board_register("rstudio")
```

And then create pins as usual,

```{r}
pins::pin(world_news, "worldnews", board = "rstudio")
```

We can then access this pin with ease using a relative path [worldnews/](worldnews/). You can then copy-paste the URL and retrieve the pin in your own R session using `pins`:

```{r eval=FALSE}
pins::pin_get("https://beta.rstudioconnect.com/content/5390/worldnews", board = "rstudio")
```

You can also skip the parts from the URL which the `pins` package can guess for you:

```{r eval=FALSE}
pins::pin_get("5390/worldnews", board = "rstudio")
```
