---
title: "metadata"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pins)
```

There are two versions of pins metadata:

* version 0: used in pins 0.4.5 and earlier
* version 1: used in pins 1.0.0 and later (including 0.99.9000)

Metadata versions are backward compatible, e.g. pins 1.0.0 can read versions 0 and 1. pins 1.0.0 and later will automatically throw an error recommending that you upgrade if it encounters a new metadata version.

In version 1 and greater you can identify the metadata version by consulting the "api_version" key. If it is absent, you can assume that you have version 0. `read_meta()` adds this automatically, and gives an informative error if you're reading a newer version than what is supported.

There are two major differences between version 0 and version 1:

* In version 0, type refers to the type of object (e.g. `table`, `default`, `files`).
  In version 1, type refers to the storage mechanism (e.g. `arrow`, `csv`, `rds`, `pickle`).

* In version 0, user supplied metadata is intermingled with pins metadata.
  In version 1, user supplied metadata is stored under a `user` key.
