---
title: "api"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pins)
```

# Big picture

Key concepts:

* pin version: one or more files accompanied by metadata
* pin: a collection of pin versions 
* pin board: a collection of pins

These a realised differently for different backends:

* Folder: each pin is a directory; each version is a subdirectory. Version id is content hash.
* RSC: each pin is a content item; each version is a bundle. Version id is bundle ids
* Github: each pin is a directory; each version is commit. Version id is sha1. 

## Main API

### v1.0.0

* `pin_list()`
* `pin_find()`
* `pin_browse()`
* `pin_meta()`
* `board_pin_upload()` powers `pin_upload()` and `pin_write()`. 
* `pin_cache()` powers `pin_read()` and `pin_download()`

Still not sure how to connect user facing `pin_write()`/`pin_read()` docs with underlying `pin_store()`/`pin_cache()` generics.

### Transitional

These are from v0, but are close enough to v1 semantics that I haven't touched yet.

* `board_pin_versions()`
* `board_pin_remove()`

Probably should define `pin_versions()` and `pin_delete()` and have default method fall back.

## Metadata

There are two versions of pins metadata:

* version 0: used in pins 0.4.5 and earlier
* version 1: used in pins 1.0.0 and later (including 0.99.9000)

Metadata versions are backward compatible, e.g. pins 1.0.0 can read versions 0 and 1. pins 1.0.0 and later will automatically throw an error recommending that you upgrade if it encounters a new metadata version.

In version 1 and greater you can identify the metadata version by consulting the "api_version" key. If it is absent, you can assume that you have version 0. `read_meta()` adds this automatically, and gives an informative error if you're reading a newer version than what is supported.

There are two major differences between version 0 and version 1:

* In version 0, type refers to the type of object (e.g. `table`, `default`, `files`).
  In version 1, type refers to the storage mechanism (e.g. `arrow`, `csv`, `rds`, `pickle`).

* In version 0, user supplied metadata is intermingled with pins metadata.
  In version 1, user supplied metadata is stored under a `user` key.

Metadata is built up progressively.
