---
title: "api"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pins)
```

Key concepts
* pin version: one or more files accompanied by metadata
* pin: a collection of pin versions 
* pin board: a collection of pins

(an unversioned pin deletes the old version when adding a new version)

For a file system backed board:
* a pin version is a directory containing `data.txt` + other files
* a pin is a directory of pin-versions
* a pin-board is a directory of pins

For RSC:
* a pin-version is a bundle
* a pin is a content item
* a pin-board is an RSC instance

Remote pins also manage a cache, organised around the same principles.

Primitives:

* `pin_list(board)` -> data frame.
* `pin_upload(board, path, name, metadata = list())` -> board (invisibly)
* `pin_download(board, name, version = NULL)` -> path(s)
* `pin_delete(board, name)` -> board (invisibly)
* `pin_metadata(board, name, version = NULL)` -> list
* `pin_versions(board, name)` -> data frame

Internals? Maybe need something higher level for version clean up?
* `pin_version_list()` 
* `pin_version_delete()`
* `pin_version_find()`
* `pin_data_download()`

User facing:

* `pin_search()` - `pin_ls()` + `pin_metadata()` + grepl -> data frame
* `pin_read()` - `pin_download()` + type specific behaviour -> board (invisibly)
* `pin_write()` - `pin_upload()` + type specific behaviour -> object

May need to customise via `pin_read()`/`pin_write()`, so that e.g. `board_connect()` can clearly document that you can add a `tags` parameter. But these would just be thin wrappers over `pin_download()`/`pin_upload()`.

* Folder: each pin is a directory; each version is a subdirectory. Version id is content hash.
* RSC: each pin is a content item; each version is a bundle. Version id is bundle ids
* Github: each pin is a directory; each version is commit. Version id is sha1. 


